Annotation exploration for SCPCP000006, `predicted.score` threshold `r params$predicted.score_thr`

### 2024-10-28 analyses/cell-type-wilms-tumor-06/notebook/04_annotation_Across_Samples_exploration.Rmd
update mapping_score --> predicted.score

---
title: "Annotation exploration for SCPCP000006, mapping score threshold `r params$mapping_score_thr`"
author: "Maud PLASCHKA"
date: "`r Sys.Date()`"
params: 
  mapping_score_thr: 0.85
output: 
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
---

@@

We also looked into the mapping scores for each of the compartments and explored the potential effects of score thresholds.
In this notebook, we fixed the mapping score threshold to `r params$mapping_score_thr`.

@@

      # Create and return a data frame from the Seurat object with relevant annotations
      # this data frame will have six columns: barcode, compartment, organ, compartment_score, VWF and PTPRC
      data.frame(
        compartment = srat$fetal_kidney_predicted.compartment,
        compartment_score = srat$fetal_kidney_predicted.compartment.score,
        organ = srat$fetal_full_predicted.organ, 
        PTPRC = FetchData(object = srat, vars = "ENSG00000081237", layer = "counts"),
        VWF = FetchData(object = srat, vars = "ENSG00000110799", layer = "counts"),
        umap = srat@reductions$umap@cell.embeddings
        ) |>
        tibble::rownames_to_column("barcode") |>
        dplyr::mutate(sample_id = sample_id)
    }
  ) |>
  # now combine all dataframes to make one big one
  dplyr::bind_rows() |>
  # Combine with metadata
  dplyr::left_join(metadata, by = c("sample_id" = "scpca_sample_id")) |>
  # Create column for whether the compartment score passes threshold
  dplyr::mutate(pass_mapping_QC = compartment_score > params$mapping_score_thr)

# we create a data frames of annotation of cells that pass the mapping score threshold
cell_type_df_pass <- cell_type_df  |>
  dplyr::filter(pass_mapping_QC)

# we create a data frames of annotation of cells that don't pass the mapping score threshold
cell_type_df_nopass <- cell_type_df  |>
  dplyr::filter(!pass_mapping_QC)
```

@@

DT::datatable(compartment_df, 
              caption = "counts of cells in each compartment",
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',
                               buttons = c( 'csv', 'excel')))
```
### Label transfer (mapping) scores for the four compartments

@@

```{r fig.height=5, fig.width=10, message=FALSE, warning=FALSE, out.width='100%'}
p <- do_Feature_densityplot( df = cell_type_df,
                    feature = "compartment_score",
                    group.by = "compartment"
                    )
p +  geom_vline(xintercept=params$mapping_score_thr)
```

Some cells have an excellent mapping score (> `r params$mapping_score_thr`) while other performed worse. Let's have a look at the compartments per patient.

@@

p + geom_hline(yintercept = params$mapping_score_thr)  + coord_cartesian(ylim = c(0.25,1))

@@

Regarding the endothelial cells, some patient have very high mapping scores while others perform poorly. 
Let's have a look if this can be linked to some metadata/clinical data.

@@

We finally look for each patient and compartment, the number of cells that have a mapping score > `r params$mapping_score_thr` (TRUE).

@@

```{r fig.height=6, fig.width=15, message=FALSE, warning=FALSE, out.width='100%'}
do_Feature_mean( df = cell_type_df_pass,
                    feature = "ENSG00000110799",
                    group.by = "compartment") +
  ggtitle("VWF expression averaged by compartment for cells passing the mapping score threshold")
```

```{r fig.height=6, fig.width=15, message=FALSE, warning=FALSE, out.width='100%'}
do_Feature_mean( df = cell_type_df_nopass,
                    feature = "ENSG00000110799",
                    group.by = "compartment") +
  ggtitle("VWF expression averaged by compartment for cells that don't pass the mapping score threshold")
```

@@

```{r fig.height=6, fig.width=15, message=FALSE, warning=FALSE, out.width='100%'}
do_Feature_mean( df = cell_type_df_pass,
                    feature = "ENSG00000081237",
                    group.by = "compartment") +
  ggtitle("PTPRC expression averaged by compartment for cells passing the mapping score threshold")
```

```{r fig.height=6, fig.width=15, message=FALSE, warning=FALSE, out.width='100%'}
do_Feature_mean( df = cell_type_df_nopass,
                    feature = "ENSG00000081237",
                    group.by = "compartment") +
  ggtitle("PTPRC expression averaged by compartment for cellsthat are not passing the mapping score threshold")
```

@@

Black cross show cell that do not pass the mapping threshold.

@@

#### Cells passing the mapping score threshold

Point are colored per compartment. 
Here we only plotted cells that do pass the mapping score threshold.

>>>>

---
title: "Annotation exploration for SCPCP000006, `predicted.score` threshold `r params$predicted.score_thr`"
author: "Maud PLASCHKA"
date: "`r Sys.Date()`"
params: 
  predicted.score_thr: 0.85
output: 
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
---

@@

We also looked into the `predicted.score` for each of the compartments and explored the potential effects of score thresholds.
In this notebook, we fixed the `predicted.score` threshold to `r params$predicted.score_thr`.

@@

      # Create and return a data frame from the Seurat object with relevant annotations
      # this data frame will have six columns: barcode, compartment, organ, compartment_score, VWF and PTPRC
      data.frame(
        compartment = srat$fetal_kidney_predicted.compartment,
        compartment_score = srat$fetal_kidney_predicted.compartment.score,
        organ = srat$fetal_full_predicted.organ, 
        PTPRC = FetchData(object = srat, vars = "ENSG00000081237", layer = "counts"),
        VWF = FetchData(object = srat, vars = "ENSG00000110799", layer = "counts"),
        umap = srat@reductions$umap@cell.embeddings
        ) |>
        tibble::rownames_to_column("barcode") |>
        dplyr::mutate(sample_id = sample_id)
    }
  ) |>
  # now combine all dataframes to make one big one
  dplyr::bind_rows() |>
  # Combine with metadata
  dplyr::left_join(metadata, by = c("sample_id" = "scpca_sample_id")) |>
  # Create column for whether the compartment score passes threshold
  dplyr::mutate(pass_mapping_QC = compartment_score > params$predicted.score_thr)

# we create a data frames of annotation of cells that pass the `predicted.score` threshold
cell_type_df_pass <- cell_type_df  |>
  dplyr::filter(pass_mapping_QC)

# we create a data frames of annotation of cells that don't pass the `predicted.score` threshold
cell_type_df_nopass <- cell_type_df  |>
  dplyr::filter(!pass_mapping_QC)
```

@@

### Label transfer predicted.score for the four compartments

@@

```{r fig.height=5, fig.width=10, message=FALSE, warning=FALSE, out.width='100%'}
p <- do_Feature_densityplot( df = cell_type_df,
                    feature = "compartment_score",
                    group.by = "compartment"
                    )
p +  geom_vline(xintercept=params$predicted.score_thr)
```

Some cells have an excellent `predicted.score` (> `r params$predicted.score_thr`) while other performed worse. Let's have a look at the compartments per patient.

@@

p + geom_hline(yintercept = params$predicted.score_thr)  + coord_cartesian(ylim = c(0.25,1))

@@

Regarding the endothelial cells, some patient have very high `predicted.score` while others perform poorly. 
Let's have a look if this can be linked to some metadata/clinical data.

@@

We finally look for each patient and compartment, the number of cells that have a `predicted.score` > `r params$predicted.score_thr` (TRUE).

@@

```{r fig.height=6, fig.width=15, message=FALSE, warning=FALSE, out.width='100%'}
do_Feature_mean( df = cell_type_df_pass,
                    feature = "ENSG00000110799",
                    group.by = "compartment") +
  ggtitle("VWF expression averaged by compartment for cells passing the `predicted.score` threshold")
```

```{r fig.height=6, fig.width=15, message=FALSE, warning=FALSE, out.width='100%'}
do_Feature_mean( df = cell_type_df_nopass,
                    feature = "ENSG00000110799",
                    group.by = "compartment") +
  ggtitle("VWF expression averaged by compartment for cells that don't pass the `predicted.score` threshold")
```

@@

```{r fig.height=6, fig.width=15, message=FALSE, warning=FALSE, out.width='100%'}
do_Feature_mean( df = cell_type_df_pass,
                    feature = "ENSG00000081237",
                    group.by = "compartment") +
  ggtitle("PTPRC expression averaged by compartment for cells passing the `predicted.score` threshold")
```

```{r fig.height=6, fig.width=15, message=FALSE, warning=FALSE, out.width='100%'}
do_Feature_mean( df = cell_type_df_nopass,
                    feature = "ENSG00000081237",
                    group.by = "compartment") +
  ggtitle("PTPRC expression averaged by compartment for cells that are not passing the predicted score threshold")
```

@@

Black cross show cell that do not pass the `predicted.score` threshold.

@@

#### Cells passing the `predicted.score` threshold

Point are colored per compartment. 
Here we only plotted cells that do pass the `predicted.score` threshold.
